\thetitle

### 2020-11-10 sun_phases/spin_qudit_tomo.tex
renamed some equations and fixed an error in Eq. B32

Equipped with the bound on variances in Eq.~\eqref{eq:variance_bound}, we again consider the block-diagonalized form of the measurement matrix derived in Section \ref{sec:bound}:
\begin{align}
  U M_V = \sum_\ell \op{\ell} \otimes M_{V\ell},
  &&
  M_{V\ell} = \sum_{m,\v v} D^\ell_{m,0} \op{m}{\v v}.
\end{align}
The block diagonal structure of this matrix implies that, rather than an integer index $k\in\set{1,2,\cdots,d^2}$, we can identify the singular values and right singular vectors of $UM_V$ (and equivalently $M_V$) by a block index $\ell\in\set{0,1,\cdots,d-1}$ and a sub-index $m=\set{-\ell,-\ell+1,\cdots,\ell}$.
Making the replacement $k\to\p{\ell,m}$, we can thus repeat the derivations through Eq.~\eqref{eq:bound_eps} of Section \ref{sec:bound} without the assumption of Eq.~\eqref{eq:error_assumption}, finding that
\begin{align}
  \Bbk{\abs{\tilde\rho_{\ell m}^V-\rho_{\ell m}^V}^2}
  = \sum_{i,i'} q_{\ell m i}^V \p{q_{\ell m i'}^V}^*
  \bbk{\epsilon_{i\ell} \epsilon_{i'\ell}}
  = \sum_i \abs{q_{\ell m i}^V}^2
  \bbk{\epsilon_{i\ell}^2}
  \le \f{2\ell+1}{nd} \times \gamma_\ell^2
  \times \p{1-\delta_{\ell,0}},
  \label{eq:bound_single}
\end{align}
where $i$ indexes an axis $\v v_i\in V$.
Altogether, we thus find that $n$ measurements of spin projection along every axis $\v v\in V$ allows us to construct an empirical estimate $\tilde\rho_V$ of an unknown state $\rho$ that is accurate to within a mean squared distance
\begin{align}
  \Bbk{\norm{\tilde\rho_V-\rho}^2}
  = \sum_{\ell,m} \Bbk{\abs{\tilde\rho_{\ell m}^V-\rho_{\ell m}^V}^2}
  < \f1{nd} \sum_{\ell>0} \p{2\ell+1} \gamma_\ell^2 \S_{V\ell}^2.
  \label{eq:bound_full}
\end{align}
We note that whereas the bound on $\bbk{\epsilon_{i\ell}^2}$ in Eq.~\eqref{eq:variance_bound} is tight, the bounds in Eq.~\eqref{eq:bound_single} and Eq.~\eqref{eq:bound_full} are not: the upper bound on $\bbk{\epsilon_{i\ell}^2}$ cannot be achieved simultaneously for all axes $\v v_i$ or degrees $\ell$.
In principle, there is therefore room for improvement on the bound in Eq.~\eqref{eq:bound_full}.

As a final comment, we note that the statistical error bound in \eqref{eq:bound_full} is a ``worst-case'' bound that is agnostic to the state $\rho$.
The actual statistical error for any particular state may be considerably smaller, and can be obtained by substituting the exact expression for $\bbk{\epsilon_{i\ell}^2}$:
\begin{align}
  \Bbk{\norm{\tilde\rho_V-\rho}^2}
  = \sum_{i,\ell,m} \abs{q^V_{\ell mi}}^2 \bbk{\epsilon_{i\ell}^2}
  = \f1n \sum_{i,\ell} C^V_{\ell i} \sigma_\rho^2\p{T_{i\ell}},
  &&
  C^V_{\ell i} \equiv \sum_m \abs{q_{\ell mi}}^2,
  \label{eq:error_full_start}
\end{align}
where $\sigma_\rho^2\p{X}\equiv\bk{X^2}_\rho - \bk{X}_\rho^2$ is the variance of $X$.
To simplify this error, we use the fact that all $T_{i\ell}=T_{i\ell}^\dag$ to expand
\begin{align}
  \sigma_\rho^2\p{T_{i\ell}}
  = \sigma_\rho^2\p{\sum_m D^\ell_{m,0}\p{\v v_i} T_{\ell m}}
  = \sum_{m,n} D^\ell_{m,0}\p{\v v_i}^* D^\ell_{n,0}\p{\v v_i}
  \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}},
\end{align}
where
\begin{align}
  \cov_\rho\p{X,Y} \equiv \bk{X Y}_\rho - \bk{X}_\rho \bk{Y}_\rho
\end{align}
is the covariance between $X$ and $Y$.
This expansion allows us to express the error in Eq.~\eqref{eq:error_full_start} more compactly as
\begin{align}
  \Bbk{\norm{\tilde\rho_V-\rho}^2}
  = \f1n \sum_\ell \p{\xi_{V\ell}|\T_{\rho\ell}},
  \label{eq:error_full}
\end{align}
where
\begin{align}
  \xi_{V\ell} \equiv M_{V\ell} C_{V\ell} M_{V\ell}^\dag,
  &&
  C_{V\ell}
  \equiv \diag\p{C^V_{\ell,1},C^V_{\ell,2},\cdots,C^V_{\ell,\abs{V}}}
\end{align}
and
\begin{align}
  \T_{\rho\ell}
  \equiv \sum_{m,n} \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}} \op{m}{n}
  = \sum_{m,n} \sp{\delta_{mn} \bk{T_{\ell m}^\dag T_{\ell m}}_\rho
    - \bk{T_{\ell m}^\dag}_\rho \bk{T_{\ell n}}_\rho} \op{m}{n}.
\end{align}
While the error in Eq.~\eqref{eq:error_full} cannot be known exactly without knowing $\rho$, this error can be estimated by computing the right-hand side of Eq.~\eqref{eq:error_full} with the replacement $\rho\to\tilde\rho_V$.

>>>>

Equipped with the bound on variances in Eq.~\eqref{eq:variance_bound}, we again consider the block-diagonalized form of the measurement matrix derived in Section \ref{sec:bound}:
\begin{align}
  U M_V = \sum_\ell \op{\ell} \otimes M_{V\ell},
  &&
  M_{V\ell} = \sum_{m,\v v} D^\ell_{m,0} \op{m}{\v v}.
\end{align}
The block diagonal structure of this matrix implies that, rather than an integer index $k\in\set{1,2,\cdots,d^2}$, we can identify the singular values and right singular vectors of $UM_V$ (and equivalently $M_V$) by a block index $\ell\in\set{0,1,\cdots,d-1}$ and a sub-index $m=\set{-\ell,-\ell+1,\cdots,\ell}$.
Making the replacement $k\to\p{\ell,m}$, we can thus repeat the derivations through Eq.~\eqref{eq:bound_eps} of Section \ref{sec:bound} without the assumption of Eq.~\eqref{eq:error_assumption}, finding that
\begin{align}
  \Bbk{\abs{\tilde\rho_{\ell m}^V-\rho_{\ell m}^V}^2}
  = \sum_{i,i'} q_{\ell m i}^V \p{q_{\ell m i'}^V}^*
  \bbk{\epsilon_{i\ell} \epsilon_{i'\ell}}
  = \sum_i \abs{q_{\ell m i}^V}^2
  \bbk{\epsilon_{i\ell}^2}
  \le \f{2\ell+1}{nd} \times \gamma_\ell^2
  \times \p{1-\delta_{\ell,0}},
  \label{eq:bound_single}
\end{align}
where $i$ indexes an axis $\v v_i\in V$.
Altogether, we thus find that $n$ measurements of spin projection along every axis $\v v\in V$ allows us to construct an empirical estimate $\tilde\rho_V$ of an unknown state $\rho$ that is accurate to within a mean squared distance
\begin{align}
  \Bbk{\norm{\tilde\rho_V-\rho}^2}
  = \sum_{\ell,m} \Bbk{\abs{\tilde\rho_{\ell m}^V-\rho_{\ell m}^V}^2}
  < \f1{nd} \sum_{\ell>0} \p{2\ell+1} \gamma_\ell^2 \S_{V\ell}^2.
  \label{eq:bound_SM}
\end{align}
We note that whereas the bound on $\bbk{\epsilon_{i\ell}^2}$ in Eq.~\eqref{eq:variance_bound} is tight, the bounds in Eq.~\eqref{eq:bound_single} and Eq.~\eqref{eq:bound_SM} are not: the upper bound on $\bbk{\epsilon_{i\ell}^2}$ cannot be achieved simultaneously for all axes $\v v_i$ or degrees $\ell$.
In principle, there is therefore room for improvement on the bound in Eq.~\eqref{eq:bound_SM}.

As a final comment, we note that the statistical error bound in \eqref{eq:bound_SM} is a ``worst-case'' bound that is agnostic to the state $\rho$.
The actual statistical error for any particular state may be considerably smaller, and can be obtained by substituting the exact expression for $\bbk{\epsilon_{i\ell}^2}$:
\begin{align}
  \Bbk{\norm{\tilde\rho_V-\rho}^2}
  = \sum_{i,\ell,m} \abs{q^V_{\ell mi}}^2 \bbk{\epsilon_{i\ell}^2}
  = \f1n \sum_{i,\ell} C^V_{\ell i} \sigma_\rho^2\p{T_{i\ell}},
  &&
  C^V_{\ell i} \equiv \sum_m \abs{q_{\ell mi}}^2,
  \label{eq:error_SM_start}
\end{align}
where $\sigma_\rho^2\p{X}\equiv\bk{X^2}_\rho - \bk{X}_\rho^2$ is the variance of $X$.
To simplify this error, we use the fact that all $T_{i\ell}=T_{i\ell}^\dag$ to expand
\begin{align}
  \sigma_\rho^2\p{T_{i\ell}}
  = \sigma_\rho^2\p{\sum_m D^\ell_{m,0}\p{\v v_i} T_{\ell m}}
  = \sum_{m,n} D^\ell_{m,0}\p{\v v_i}^* D^\ell_{n,0}\p{\v v_i}
  \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}},
\end{align}
where
\begin{align}
  \cov_\rho\p{X,Y} \equiv \bk{X Y}_\rho - \bk{X}_\rho \bk{Y}_\rho
\end{align}
is the covariance between $X$ and $Y$.
This expansion allows us to express the error in Eq.~\eqref{eq:error_SM_start} more compactly as
\begin{align}
  \Bbk{\norm{\tilde\rho_V-\rho}^2}
  = \f1n \sum_\ell \p{\xi_{V\ell}|\T_{\rho\ell}},
  \label{eq:error_SM}
\end{align}
where
\begin{align}
  \xi_{V\ell} \equiv M_{V\ell} C_{V\ell} M_{V\ell}^\dag,
  &&
  C_{V\ell}
  \equiv \diag\p{C^V_{\ell,1},C^V_{\ell,2},\cdots,C^V_{\ell,\abs{V}}},
  &&
  \T_{\rho\ell}
  \equiv \sum_{m,n} \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}} \op{m}{n}.
\end{align}
While the error in Eq.~\eqref{eq:error_SM} cannot be known exactly without knowing $\rho$, this error can be estimated by computing the right-hand side of Eq.~\eqref{eq:error_SM} with the replacement $\rho\to\tilde\rho_V$.
